---
title: "Report Title (.)"
subtitle: "https://github.com/yangzi33/sta304-ps3"
author: 
    - Boyu Cao
    - Jiayi Yu
    - Yijia Liu
    - Ziyue Yang
date: "15/10/2020"
output: pdf_document
abstract: "This is an abstract. (.)"
---


```{r setup, include=FALSE, cache=FALSE}
# Setting seed as the last four digits of Ziyue Yang's student number
set.seed(4759)
library(tidyverse)
library(brms)
# Run modeling
source('model.R')

# Configuring global figure settings
knitr::opts_chunk$set(fig.height=3, fig.width=6, echo = FALSE)

# df <- read.csv('gss-time-stress/gss-time-stress.csv')
# df

```

# Introduction (.)

A topic that intrigues us is the relationship of marital status to subjective well-being of individuals. One of the measurement of well-being is reported mental health, which will the focus our analyses. 

*In the modern society, there exists various issues. (blah blah blah)*

# Data (.)

Throughout the article, regression models will be illustrated using data from the twenty fourth cycle of the  **General Social Survey (GSS)** on time-stress and well-being, collected in 2010^[1]^. 

The target population for this survey includes all persons 15 years of age and older in Canada, excluding residents of the Yukon, Northwest Territories, and Nunavut, as well as full-time residents of institutions. Computer assisted telephone interviewing was used to collect data for GSS, and households were reached by calling a series of randomly-generated phone numbers.

In the GSS, mental health is assessed by a question asking "In general, would you say your mental health is: ...", with provided answers "Excellent, Very Good, Good, Fair", or "Don't know" for those who are unsure about answering. Of the total of 15,390 respondents in the survey, 203 did not choose to answer, and 45 answered "Don't know". Therefore, analyses in our report are based on 15,142 respondents providing valid answers to this question. 

Due to the fact that there are merely four values, we treat them as a binary variable, by coding value $1$ for those who answered "Excellent" and "Very Good", and $0$ otherwise. The mean of this binary variable is 0.608 or 7224/11873, the proportion of those who are "mentally healthy" in the sample, denoted $p$, which can be viewed as the probability that a randomly selected sample will be "mentally healthy".

Our focus will be modeling mental health as a function of various variables. Two of the independent variables are treated as interval: age, income, while gender and marital status will be dummies. We will examine the relationship between mental health and marital status.

<!-- - Marital status: single, married, living common-law, divorced, and widowed. This will be the main response variable -->
<!-- - Age group of the respondent: by groups of 10 such as "45 to 54". -->
<!-- - Gender of the respondent: male and female. -->
<!-- - Annual income grouped by ranges:  -->
<!--   - No income,  -->
<!--   - Less than \$5000, -->
<!--   - 7788 -->
<!--   - \$0 to \$49,999, \$50,000 to \$100,000, over \$100,000. -->


```{r, echo=FALSE}
raw_plot <- ggplot(df, aes(x = marital, y = mental_dummy)) + geom_point()
raw_plot
```

*For each denoted $X_1,X_2,\dots,X_k$, .. respectively ... Consider ...*

**Frame, Plot, Raw Plot**


# Model (.)


# Results (.)


# Discussion (.)

## Weaknesses and Potential Improvements


# References (.)


<!-- For brms -->
<!-- ######## -->

* [gss]: General Social Survey (GSS) on Family (cycle 31), 2017

* Bürkner P. C. (2017). brms: An R Package for Bayesian Multilevel Models using Stan. Journal of Statistical Software. 80(1), 1-28. doi.org/10.18637/jss.v080.i01

* Bürkner P. C. (2018). Advanced Bayesian Multilevel Modeling with the R Package brms. The R Journal. 10(1), 395-411. doi.org/10.32614/RJ-2018-017

<!-- ######## -->

<!-- For Stan -->
<!-- ######## -->

* Carpenter B., Gelman A., Hoffman M. D., Lee D., Goodrich B., Betancourt M., Brubaker M., Guo J., Li P., and Riddell A. (2017). Stan: A probabilistic programming language. Journal of Statistical Software. 76(1). 10.18637/jss.v076.i01

<!-- ######## -->

* R Core Team (2020). R: A language and environment for statistical computing. R Foundation for Statistical
  Computing, Vienna, Austria. URL https://www.R-project.org/.
  
* Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source Software, 4(43), 1686,
  https://doi.org/10.21105/joss.01686

* Sam Firke (2020). janitor: Simple Tools for Examining and Cleaning Dirty Data. R package version 2.0.1.
  https://CRAN.R-project.org/package=janitor

# Appendix (. Optional)